---
title: "Tidy data"
author: "Ashwin Malshe"
date: "9/8/2020"
output:
  html_document:
    theme: sandstone
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(printr) # This package prints help in the rmarkdown document after knitting
```

For this exercise we will use `mpg` dataset, which is bundled with `ggplot2`. You don't have to explicitly call `ggplot2` into this R session because we loaded it already using `tidyverse`.

## `mpg` dataset

First, let's understand what is inside this dataset. A `?` followed by the name of the object/function/package will lead to opening up the help menu on right.

```{r}
?mpg # OR help(mpg)
```


### Exercise 1: Explore

In this exercise, we will explore the dataset. 

`summary()` will output simple summary for all the variables in the dataset

```{r}
summary(mpg)
```

Let's look at the frequency of character variables using `table()` function

```{r}
table(mpg$manufacturer)
```


`table()` outputs a vector of frequencies. What if you wanted a dataframe instead? We will use `count()` function from `dplyr`

```{r}
count(mpg, manufacturer)

```

It will work on numeric variables as well.

```{r}
count(mpg, year)
```


#### Printing specific statistics

Use these functions to calculate summary statistics for `hwy` variable.

```{r eval = FALSE}
mean()
median()
sd()
min()
max()
```


### Create a dataframe with the summary statistics of `hwy`

*Here also introduce `pipe` operator ` %>% `*


```{r}
mpg_sum1 <- mpg %>% 
  summarize(hwy_mean = mean(hwy),
         hwy_med = median(hwy),
         hwy_sd = sd(hwy))
```


```{r}
print(mpg_sum1)
```


### Grouping and summarizing

A common operation in data prep is to group datasets by a variable and then summarize the variable.

For instance, above we got summary statistics for all the full data set. What if we wanted it separately for 1999 and 2008?

```{r}
mpg_sum2 <- mpg %>% 
  group_by(year) %>% 
    summarize(hwy_mean = mean(hwy),
           hwy_med = median(hwy),
           hwy_sd = sd(hwy)) %>% 
  ungroup()
```


```{r}
print(mpg_sum2)
```

*Experimental feature using `.groups`*

```{r}
mpg_sum3 <- mpg %>% 
  group_by(year) %>% 
    summarize(hwy_mean = mean(hwy),
           hwy_med = median(hwy),
           hwy_sd = sd(hwy),
           .groups = "drop") 
```

```{r}
print(mpg_sum3)
```


## Merging dataframes

The simplest task is to attach datadrames by columns or rows without paying any consideration for a common variable to match on. For this, R has two functions:

### `rbind`

If we want to put dataframes on top of each other, we can use `rbind()`

For `rbind()` to work, the two dataframes should have:

1. The same number of columns
2. The same column names
3. The same `class` of columns

The easiest demo for this can be created by splitting an already existing dataframe.
Let's separate out `mpg` by year and then merge the two resulting dataframes back

```{r}
mpg_1999 <- mpg %>%  filter(year == 1999)
mpg_2008 <- mpg %>%  filter(year == 2008)
```

```{r}
mpg_rows <- rbind(mpg_1999, mpg_2008)
```

```{r}
summary(mpg_rows)
```



### `cbind`

If you want to merge two dataframes sideways, you can use `cbind()`

For it to work,

1. The two dataframes should have the same number of rows.

Now, let's first split `mpg` along the columns. First dataframe will have first 5 columns and the second dataframe will hold the remaining columns

```{r}
names(mpg)
```


```{r}

mpg_left <- mpg %>% select(1:5)
mpg_right <- mpg %>% select(-c(1:5))

```



```{r}
mpg_cols <- cbind(mpg_left, mpg_right)
```

```{r}
summary(mpg_cols)
```


### `*_join()` functions from `dplyr`

Commonly, you need to merge datasets sideways by adding columns. However, there are two common issues:

1. The two datasets may not have the same number of rows
2. Even if the two datasets have the same number of rows, there is no guranatee that you are matching the rows correctly.


Let's understand the second issue first.
Sort `mpg_right` by `cty`

```{r}
mpg_right2 <- mpg_right %>% 
  arrange(cty)

```

Now if you use `cbind()` on `mpg_left` and `mpg_right2` you won't get back the same original dataframe. This is because, we have mismatched the observations.

How can we resolve this issue?

```{r}

```

### Merge summary back to the original dataframe

The common key helps even when we don't have the same number of observations in the two dataframes


```{r}
mpg2 <- mpg %>% 
  inner_join(mpg_sum2, by = "year")
```










