---
title: "Homework 1"
author: "Victoria Espinola"
date: "`r sys.Date()"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
theme_set(theme_minimal())

# Set your theme here
# Create a duplicate copy of diamonds data here

```



In this homework, you will use `diamonds` data set from `ggplot2` package. It is automatically loaded when you execute `library(ggplot2)` so you don't have to separately load it. Please create a duplicate of `diamonds` and use that for homework. This will avoid corrupting the original data set.

Make sure that you understand the variables in the data by checking the help file by executing `help(diamonds)` in the console. The data has five variables that measure the dimensions. 
```{r}
myset <- diamonds
```

## Instructions

1. You are expected to recreate each plot **exactly** as shown in this homework.
2. The objective of this homework is to help you develop fine grain control over `ggplot2`. As such, please refrain from exercising artistic freedom!
3. All the plots use `theme_minimal()`. You can set this as your default theme by adding this line in the `setup` chunk **after** you load `ggplot2` library:
`theme_set(theme_minimal())`


## Q1 (3 points)

Recreate the following graph. The parameter that controls transparency is set at 0.3. You need not get exactly the same colors but they must be **discrete** and should not follow a color gradient.


```{r Q1}

ggplot(myset, aes(carat, price, fill=as.factor(clarity)))+
  geom_point(size=2, shape=21, alpha=0.3)+
   labs(x= "Diamond Carats", y= "Diamond Price", 
            title= "Scatterplot of Diamond Prices")+
  scale_y_continuous(labels = scales::dollar_format(prefix = "$"))
  
  

```


## Q2 (2 points)

The previous graph looks cluttered. So you decided to use facets instead. Recreate the following graph:


```{r Q2}
ggplot(myset, aes(carat, price, col=as.character(clarity)))+
  geom_point(alpha=0.3, fill=NA)+
  facet_wrap(vars(clarity), 
             nrow=2, 
              ncol=4)+
   labs(x= "Diamond Carats", y= "Diamond Price",
            title= "Scatterplot of Diamond Prices")+
  scale_y_continuous(labels = scales::dollar_format(prefix = "$"))+
  theme(legend.title= element_blank(), legend.position = "none")

```

##Q3 (5 points)

Next, you want to know whether the price of diamonds depends on table and depth. Note the line types. Recreate the following graphs:


```{r Q3.1, message=FALSE}
ggplot(myset, aes(table, price)) +
  stat_smooth(method="lm",se=TRUE, color="red", linetype="dashed")+
  xlim(0, 100)
  
  
   
  
```


```{r Q3.2, message=FALSE}
ggplot(myset, aes(depth, price)) +
  stat_smooth(method="lm",se=TRUE, color="white", linetype="dotdash")+
  xlim(0, 80)
  
   
```



## Q4 (5 points)

Recreate each of the following graphs for data exploration:


```{r Q5.1, message=FALSE}
ggplot(myset, aes(x*y*z, price,col=cut))+
  geom_point()+
  scale_color_manual(values=c("#99B898","#FECEAB", "#FF847C", "#E84A5F", "#2A363B"))
```


```{r Q5.2}
ggplot(myset, aes(x=price))+
  geom_histogram(bins=80,color="white")+
  scale_x_continuous(labels = scales::dollar_format(prefix = "$"))
  
```



```{r Q5.3}
ggplot(myset, aes(x=clarity,fill=clarity))+
  geom_bar(color="black")+
  scale_fill_manual(values=c("#4287F5", "#42F5E3", "#AAF542", "#F2F542", "#F5C542","#F55742", "#F542EC", "#B042F5"))+
  theme(legend.title= element_blank(), legend.position = "none")

```


```{r Q5.4}
ggplot(myset, aes(x=cut, y=depth))+
  geom_violin(color="blue")+
  geom_jitter(color="red", alpha=.3)
  
```


```{r Q5.5, message=FALSE}
ggplot(myset, aes(x=x, y=price))+
  geom_smooth(se=FALSE)+
  geom_smooth(method="lm", color="green")

```


